Instructions for developing new pages for the MarineOmics webpage

### 2023-12-14 DRAFT_DNA_methylation.Rmd
Update DRAFT_DNA_methylation.Rmd
added information on score_min

Within Bismark, the tutorial uses Bowtie 2 as the alignment software.

@@

@@

For aligning the trimmed reads to the genome the following code structure is used

@@

This will create BAM files (sequence alignment files)

@@

    -   **`--path_to_bowtie ${bowtie2_dir}`**: This option specifies the path to the Bowtie2 aligner, which Bismark uses for the alignment process.

@@

    -   **`-p 4`**: Sets the number of processing threads to 4, allowing Bismark to perform the alignment in parallel, which can speed up the process.

    -   **`-score_min L,0,-0.6`**: This sets the scoring threshold for alignments, impacting the sensitivity and specificity of the alignment.

@@

>>>>

Within Bismark, the tutorial uses Bowtie 2 (@https://doi.org/10.1038/nmeth.1923) as the alignment software. Although the Bismark user manual covers the necessary Bowtie 2 parameters, it may be useful to refer to the [Bowtie 2 manual](https://bowtie-bio.sourceforge.net/bowtie2/manual.shtml).

@@

This preparation step only needs to happen once for a given genome.

@@

Once the genome is prepared, reads can be aligned. For aligning the trimmed reads to the genome the following code structure is used:

@@

This will create BAM files (sequence alignment files) and txt files (mapping reports). The reports contain information such as mapping efficiency.

@@

    -   **`--path_to_bowtie ${bowtie2_dir}`**: This option specifies the path to the Bowtie 2 aligner, which Bismark uses for the alignment process.

@@

    -   **`-p 4`**: Sets the number of processing threads to 4, allowing Bismark to perform the alignment in parallel, which can speed up the process. This parameter can be changed based on your computing power.

    -   **`-score_min L,0,-0.6`**: This sets the scoring threshold for alignments, impacting the sensitivity and specificity of the alignment. The default scoring threshold used by Bismark is L,0,-0.2.

@@

### Mapping reports

EXAMPLE MAPPING REPORT FROM BISMARK AND HOW TO VISUALIZE IT

### Optimizing alignment sensitivity

The code above uses `score_min L,0,-0.6` as the scoring threshold for alignment sensitivity and specificity. This is different than the Bismark default, `score_min L,0,-0.2`. Reducing the sensitivity of the alignment can increase the percentage of reads aligned to the reference genome, but of course the quality of the alignment may decrease. A good practice is to align a subset of your data to the reference genome using different `score_min` parameters, then comparing mapping efficiency from the mapping reports. Generally, you want to choose a parameter that maximizes percent alignment and specificity. Example code

```         
find ${reads_dir}*_R1_001_val_1.fq.gz \
| xargs basename -s _R1_001_val_1.fq.gz | xargs -I{} ${bismark_dir}/bismark \
--path_to_bowtie ${bowtie2_dir} \
-genome ${genome_folder} \
-p 4 \
- u 10000 \
-score_min L,0,-0.6 \
--non_directional \
-1 ${reads_dir}{}_R1_001_val_1.fq.gz \
-2 ${reads_dir}{}_R2_001_val_2.fq.gz \
-o Mcap_tg
```

This is the same code as above, except `-u 10000` is used to run the alignment on a 10,000 bp subset of the data in each input file.

## Deduplication
